---
title: "Rasch model fit check"
author: "Christian Rubjerg Hejstvig-Larsen"
output:
  html_document:
    df_print: paged
    toc: true
    code_folding: hide
---

```{r}
knitr::opts_chunk$set(echo = TRUE, results = "hold")

# Load libraries
library(eRm)          # Rasch model

library(dplyr)        # Data manipulation
library(data.table)   # Data manipulation
library(reshape)      # Data manipulation

library(readr)        # Data import

library(ggplot2)      # Data visualization
library(RASCHplot)    # Data visualization
```



# Observational data
```{r}
# Load data
data <- read.csv("C:/Users/brf337/Desktop/Rasch package/data.csv", sep = "\t")
summary(data)
```

In this script we wish to check how model validation works in R and if it aligns with Python. We first look at an observational data set containing data collected online with an interactive version of the the Rosenberg Self-Esteem Scale. To test the model validation tools for the Rasch model we dichotomize the data:

```{r}
# Dichotomize first 10 columns of data
data_dichotomized <- data %>% 
  mutate_at(.vars = vars(1:10), .funs = ~ ifelse(. >= 3, 1, 0))
head(data_dichotomized)
```

Having prepared the data to be dichotomized we fit the Rasch model to the data.

```{r}
# Fit Rasch model
RaschFit <- RM(data_dichotomized[,1:10], sum0 = TRUE)
summary(RaschFit)
```

We proceed to check the item fit statistics:

```{r}
ppRaschFit <- person.parameter(RaschFit)
ifRaschFit <- itemfit(ppRaschFit)
ifRaschFit
```


# Simulated data

```{r}
# Simulate data with misspecified items
set.seed(03022025)
SML_sim_mis <- SML_sim(k = 10, n = 1000, manual_diffs = c(1, -2, 0.5, -1, 1.5, 2, -3, 0.5, -1, 1.5),
                       misspecified_items = c("item3", "item9"))
write.csv(SML_sim_mis, "C:/Users/brf337/Desktop/Rasch package/SML_sim_mis.csv")
```


```{r}
RaschFitMis <- RM(SML_sim_mis, sum0 = TRUE)
summary(RaschFitMis)
```


```{r}
ppRaschFitMis <- person.parameter(RaschFitMis)
ifRaschFitMis <- itemfit(ppRaschFitMis)
ifRaschFitMis
```




```{r}
par(mfrow = c(2, 5))
for (i in 1:ncol(SML_sim_mis)) {
  plotICC(RaschFitMis, item.subset = i, empICC = list("raw"), empCI = list())
}
```







# Simulation study

We run a simulation study to investegate the distribution of the Infit and Outfit statistics for the Rasch model. 

```{r}
n_person__list <- c(100, 150, 200, 250, 300, 350, 400, 600, 800, 1000, 1500, 2000)
result_list <- vector("list", length(n_person__list))

manual_diffs <- c(1, -2, 0.5, -1, 1.5, 2, -3, 0.5, -1, 1.5)

for (i in 1:length(n_person__list)) {
  print(paste("Running simulation for", n_person__list[i], "persons"))
  InfitOutfitSimStudy <- run_InfitOutfit_simulation_study(n_persons = n_person__list[i], k_items = 10, n_simulations = 1000, manual_diffs = manual_diffs)
  result_list[[i]] <- InfitOutfitSimStudy
}
```

We store the 0.025 and 0.975 of all the statistics in a data frame

```{r}
quantile_probs <- c(0.025, 0.05, 0.1, 0.9, 0.95, 0.975)
min_infit_list <- lapply(result_list, function(x) quantile(x$mininfit, probs = quantile_probs))
max_infit_list <- lapply(result_list, function(x) quantile(x$maxinfit, probs = quantile_probs))
min_outfit_list <- lapply(result_list, function(x) quantile(x$minoutfit, probs = quantile_probs))
max_outfit_list <- lapply(result_list, function(x) quantile(x$maxoutfit, probs = quantile_probs))

mininfit_values <- sapply(1:length(quantile_probs), function(i) unlist(lapply(min_infit_list, `[[`, i)))
maxinfit_values <- sapply(1:length(quantile_probs), function(i) unlist(lapply(max_infit_list, `[[`, i)))
minoutfit_values <- sapply(1:length(quantile_probs), function(i) unlist(lapply(min_outfit_list, `[[`, i)))
maxoutfit_values <- sapply(1:length(quantile_probs), function(i) unlist(lapply(max_outfit_list, `[[`, i)))

colnames(mininfit_values) <- paste0("mininfit",as.character(quantile_probs))
colnames(maxinfit_values) <- paste0("maxinfit",as.character(quantile_probs))
colnames(minoutfit_values) <- paste0("minoutfit",as.character(quantile_probs))
colnames(maxoutfit_values) <- paste0("maxoutfit",as.character(quantile_probs))

mininfit_df <- data.frame(n_persons = n_person__list, mininfit_values)
maxinfit_df <- data.frame(n_persons = n_person__list, maxinfit_values)
minoutfit_df <- data.frame(n_persons = n_person__list, minoutfit_values)
maxoutfit_df <- data.frame(n_persons = n_person__list, maxoutfit_values)
```


We then plot the data frames:

```{r}
mininfitp <- ggplot(data = mininfit_df, aes(x = n_persons, y = mininfit0.025)) +
  geom_line(aes(color = "0.025")) +
  geom_line(aes(y = mininfit0.05, color = "0.05")) +
  geom_line(aes(y = mininfit0.1, color = "0.1")) +
  geom_line(aes(y = mininfit0.9, color = "0.9")) +
  geom_line(aes(y = mininfit0.95, color = "0.95")) +
  geom_line(aes(y = mininfit0.975, color = "0.975")) +
  labs(title = "Quantiles of Min Infit MS", x = "Number of persons", y = "Min Infit MS") +
  scale_x_continuous(breaks = n_person__list[c(-2, -3)]) +
  scale_color_manual(name = "Quantile", 
                     values = c("0.025" = "red",
                                "0.05" = "orange",
                                "0.1" = "yellow",
                                "0.9" = "yellow",
                                "0.95" = "orange",
                                "0.975" = "blue")) +
  scale_shape_discrete(name = "color") +
  theme_bw() + theme(axis.text.x = element_text(angle = 55, vjust = 1, hjust=1), 
                     axis.text = element_text(size = 15),  
                     axis.title=element_text(size=14))

minoutfitp <- ggplot(data = minoutfit_df, aes(x = n_persons, y = minoutfit0.025)) +
  geom_line(aes(color = "0.025")) +
  geom_line(aes(y = minoutfit0.05, color = "0.05")) +
  geom_line(aes(y = minoutfit0.1, color = "0.1")) +
  geom_line(aes(y = minoutfit0.9, color = "0.9")) +
  geom_line(aes(y = minoutfit0.95, color = "0.95")) +
  geom_line(aes(y = minoutfit0.975, color = "0.975")) +
  labs(title = "Quantiles of Min Outfit MS", x = "Number of persons", y = "Min Outfit MS") +
  scale_x_continuous(breaks = n_person__list[c(-2, -3)]) +
  scale_color_manual(name = "Quantile", 
                     values = c("0.025" = "red",
                                "0.05" = "orange",
                                "0.1" = "yellow",
                                "0.9" = "yellow",
                                "0.95" = "orange",
                                "0.975" = "blue")) +
  scale_shape_discrete(name = "color") +
  theme_bw() + theme(axis.text.x = element_text(angle = 55, vjust = 1, hjust=1), 
                     axis.text = element_text(size = 15),  
                     axis.title=element_text(size=14))

maxinfitp <- ggplot(data = maxinfit_df, aes(x = n_persons, y = maxinfit0.025)) +
  geom_line(aes(color = "0.025")) +
  geom_line(aes(y = maxinfit0.05, color = "0.05")) +
  geom_line(aes(y = maxinfit0.1, color = "0.1")) +
  geom_line(aes(y = maxinfit0.9, color = "0.9")) +
  geom_line(aes(y = maxinfit0.95, color = "0.95")) +
  geom_line(aes(y = maxinfit0.975, color = "0.975")) +
  labs(title = "Quantiles of Max Infit MS", x = "Number of persons", y = "Max Infit MS") +
  scale_x_continuous(breaks = n_person__list[c(-2, -3)]) +
  scale_color_manual(name = "Quantile", 
                     values = c("0.025" = "red",
                                "0.05" = "orange",
                                "0.1" = "yellow",
                                "0.9" = "yellow",
                                "0.95" = "orange",
                                "0.975" = "blue")) +
  scale_shape_discrete(name = "color") +
  theme_bw() + theme(axis.text.x = element_text(angle = 55, vjust = 1, hjust=1), 
                     axis.text = element_text(size = 15),  
                     axis.title=element_text(size=14))

maxoutfitp <- ggplot(data = maxoutfit_df, aes(x = n_persons, y = maxoutfit0.025)) +
  geom_line(aes(color = "0.025")) +
  geom_line(aes(y = maxoutfit0.05, color = "0.05")) +
  geom_line(aes(y = maxoutfit0.1, color = "0.1")) +
  geom_line(aes(y = maxoutfit0.9, color = "0.9")) +
  geom_line(aes(y = maxoutfit0.95, color = "0.95")) +
  geom_line(aes(y = maxoutfit0.975, color = "0.975")) +
  labs(title = "Quantiles of Max Outfit MS", x = "Number of persons", y = "Max Outfit MS") +
  scale_x_continuous(breaks = n_person__list[c(-2, -3)]) +
  scale_color_manual(name = "Quantile", 
                     values = c("0.025" = "red",
                                "0.05" = "orange",
                                "0.1" = "yellow",
                                "0.9" = "yellow",
                                "0.95" = "orange",
                                "0.975" = "blue")) +
  scale_shape_discrete(name = "color") +
  theme_bw() + theme(axis.text.x = element_text(angle = 55, vjust = 1, hjust=1), 
                     axis.text = element_text(size = 15),  
                     axis.title=element_text(size=14))



gridExtra::grid.arrange(mininfitp, minoutfitp, maxinfitp, maxoutfitp, ncol = 2)
```

```{r}
png(file = "C:/Users/brf337/Desktop/Rasch package/R code/Rplots/infitoutfitsimR.png", width = 1100, height = 800)
gridExtra::grid.arrange(mininfitp, minoutfitp, maxinfitp, maxoutfitp, ncol = 2)
dev.off()
```







